---
title: "HW2_team"
output: html_document
date: "2024-09-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DescTools)
library(writexl)
library(tidyverse)
library(car)
library(mgcv)
library(car)
```


```{r}
bank_bin <- read.csv("insurance_t_bin.csv")
```

```{r}
#finding alpha level based on sample size
print(pchisq(log(nrow(bank_bin)),1,lower.tail = F))
# alpha level from above test is 0.002630927
```


```{r}
#finding categories with mising values
ulist <- lapply(bank_bin, unique)
ulist
#View(bank_bin)
```


For any variable with missing values, change the data to include a missing category instead of a missing value for the categorical variable. 
  - (HINT: Now all variables should be categorized (treated as categorical variables so no more continuous variable assumptions) and without missing values. Banks 
          do this for more advanced modeling purposes that we will talk about in the spring.)
          
#I found INV, CC, CCPURC, and HMOWN to have missing values that will need to be updated  

```{r}
# changing all int columns to character columns so a new missing "M" category can be added 
bank_bin1 <- bank_bin %>% mutate(across(where(is.integer), as.character))

#replacing all NA with an indicator "M"
bank_bin1 <- bank_bin1 %>% replace(is.na(.),"M")

#checking the 4 variables that had NA in original dataset
print(unique(bank_bin$INV))
print(unique(bank_bin$CC))
print(unique(bank_bin$CCPURC))
print(unique(bank_bin$HMOWN))

#checking to see if the NA's were replaced with "M" in bank_bin1
print(unique(bank_bin1$INV))
print(unique(bank_bin1$CC))
print(unique(bank_bin1$CCPURC))
print(unique(bank_bin1$HMOWN))


```

#Check each variable for separation concerns. Document in the report and adjust any variables with complete or quasi-separation concerns

```{r}
vars <- colnames(bank_bin1)

for(i in vars){
  print("---------------------------------------")
  print(i)
  print(table(bank_bin1$INS, bank_bin1[[i]]))
}
#reminder for danny --> INS will be the rows, and bank_bin[[i]] will be the columns
```

#I found 2 variables with quasi-complete seperation (and none with complete seperation)
CASHBK, MMCRED

```{r}
print(table(bank_bin1$INS, bank_bin1$CASHBK))
```

#CASHBK
(CASHBK = 2) has no occurrences of (INS = 1), so will merge (CASHBK = 2) with (CASHBK = 1) into "1+"


```{r}
#fixing CASHBK by merging (CASHBK = 1 | CASHBK = 2) into (CASHBK = "1+")
bank_bin1$CASHBK.c <- as.character(bank_bin1$CASHBK)
bank_bin1$CASHBK.c[which(bank_bin1$CASHBK %in% c("1","2"))] <- "1+"
#print(table(bank_bin1$CASHBK.c, bank_bin$CASHBK))
print(table(bank_bin1$INS, bank_bin1$CASHBK.c))
```

```{r}
print(table(bank_bin1$INS, bank_bin1$MMCRED))
```
#MMCRED

(MMCRED = 5) only has no occurrences of (INS = 1), so will merge (MMCRED = 5) with (MMCRED = 3) into "3+"
    note for danny: MMCRED = 4 does not exist in the training dataset
    Dr. Labarr --> should we also be checking in the validation set for seperation

```{r}
#fixing MMCRED by merging (MMCRED = 3 | MMCRED = 5) into (MMCRED = "3+), and using a new variable MMCRED.c 
bank_bin1$MMCRED.c <- as.character(bank_bin1$MMCRED)
bank_bin1$MMCRED.c[which(bank_bin1$MMCRED %in% c("3","5"))] <- "3+"
print(table(bank_bin1$INS, bank_bin1$MMCRED.c))
```

```{r}
#dropping CASHBK and MMCRED columns
bank_bin2 <- subset(bank_bin1, select = -c(CASHBK, MMCRED ))
#print(table(bank_bin1$INS, bank_bin1$CASHBK.c))
#print(table(bank_bin1$INS, bank_bin1$MMCRED.c))
```

```{r}
vars1 <- colnames(bank_bin1)
vars2 <- colnames(bank_bin2)

#for(i in vars2){
  #print("---------------------------------------")
  #print(i)
  #print(table(bank_bin$INS, bank_bin[[i]]))}

#making sure the two new columns were added, and they were --> 50 versus 48 ok as the old dataset had 48 original columns and then 2 new columns, 
    #and the new dataset has 46 of the original columns and then two of its columns were updated to avoid quasi-seperation (so 48 in total)
print(length(vars1))
print(length(vars2))
diff <- setdiff(vars1, vars2)
diff
```
#Build a main effects only binary logistic regression model to predict the purchase of the insurance product
```{r}
str(bank_bin2)
print(unique(bank_bin2$INS))
```

```{r}

#bank_bin1 <- bank_bin1 %>% mutate(across(where(is.character), as.factor))

#Use backward selection to do the variable selection – the Bank currently uses alpha = 0.002,and p-values to perform backward selection
full.model <- glm(factor(INS) ~ . , 
                  data = bank_bin2, family = binomial(link = "logit"))
empty.model <- glm(factor(INS) ~ 1 , 
                  data = bank_bin2, family = binomial(link = "logit"))
summary(full.model) 
```

```{r}
back.model <- step(full.model,
                  scope = list(lower = empty.model,
                               upper = full.model),
                  direction = "backward", k = qchisq(0.002, 1, lower.tail =  FALSE))

```
```{r}
summary(back.model)
```

```{r}
#summary(back.model)
#odds ration backward selection model
(exp(cbind(coef(back.model))))
```

```{r}
#investigating the issue with the CCM category having NA/singularities issues
table(bank_bin2$INV, bank_bin2$CC)
#car::vif(back.model) #--> does not run due to the following error: there are aliased coefficients in the model
alias(back.model)
```

```{r}
#trying to do equivalent of a SAS proc freq
three_way <- xtabs(~ CC + INV + INS, data=bank_bin2) 
table <- ftable(three_way)
#table
three_way
```


```{r}
#list of vars that the backwards selection model chose DDA NSF IRA INV INVM MTG CC DDABAL_BIN CHECKS_BIN TELLER_BIN
#TELLER_BIN SAVBAL_BIN ATMAMT_BIN CDBAL_BIN ILSBAL_BIN MMBAL_BIN

```


```{r}
#Forward Selection p value 0.002
full.model1 <- glm(factor(INS) ~ (DDA + NSF + IRA + INV + INV + MTG + CC + DDABAL_BIN + CHECKS_BIN  +
TELLER_BIN + SAVBAL_BIN + ATMAMT_BIN + CDBAL_BIN + ILSBAL_BIN + MMBAL_BIN)^2, data = bank_bin2, family = binomial(link = "logit"))

empty.model1 <- glm(factor(INS) ~ 1, data = bank_bin2, family = binomial(link = "logit"))

for.model <- step(empty.model1,
                  scope = list(lower = empty.model1,
                               upper = full.model1),
                  direction = "forward", k = qchisq(0.002, 1, lower.tail =  FALSE))
```


```{r}
summary(for.model)
```

```{r}
interaction1 <- xtabs(~ DDA + INS + IRA, data=bank_bin2) 
interaction1

#interaction2 <- xtabs(~ IRA + INS + DDA , data=bank_bin2) 
#interaction2
#danny does not see separation issue with interaction
```

